---
title: "Fixed effects and Random effects"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---


```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```


## Switch for party selection and spill over

```{r}
party.selec.switch <- 1
spill.over.switch <- 0 # 1: with spill over
turnout.switch <- 0 # 1: turnout analysis
```

See below what value corresponds to what party selection.

```{r}
# OLD !!!

#if (party.selec.switch==1) party.selection <- c(1:6)      # S A F L G 
#if (party.selec.switch==2) party.selection <- c(1:2,6)      # nur C und S + G (2021)
#if (party.selec.switch==3) party.selection <- c(3:6)      # ohne C und S und G (2021)
#if (party.selec.switch==4) party.selection <- c(2:6)      # ohne C
#if (party.selec.switch==5) party.selection <- c(1,3:6)    # ohne S
#if (party.selec.switch==6) party.selection <- c(1:3,5:6) # ohne F
#if (party.selec.switch==7) party.selection <- c(1:4,6)    # ohne Linke
#if (party.selec.switch==8) party.selection <- c(1:5)     # ohne G
#if (party.selec.switch==9) party.selection <- c(1:2,4:6)  # ohne AfD
#if (party.selec.switch==10) party.selection <- c(2,4:6)    # ohne C und A
#if (party.selec.switch==11) party.selection <- c(3:4,6)    # ohne C, S und L


## Do not select only one party. This won't work!
## Party positions: 1=SPD, 2=CDU/CSU, 3=GRUENE, 4=FDP, 5=AfD, 6=DIE LINKE, 7=BSW

if (party.selec.switch==1) party.selection <- c(1:7)       # All parties: S C G F A L B
if (party.selec.switch==2) party.selection <- c(1:2,3)     # SPD, CDU/CSU, GRUENE
if (party.selec.switch==3) party.selection <- c(4:7)       # FDP, AfD, DIE LINKE, BSW (without SPD, CDU/CSU, GRUENE)
if (party.selec.switch==4) party.selection <- c(1,3:7)     # Without CDU/CSU
if (party.selec.switch==5) party.selection <- c(2:7)       # Without SPD
if (party.selec.switch==6) party.selection <- c(1:3,5:7)   # Without FDP
if (party.selec.switch==7) party.selection <- c(1:5,7)     # Without DIE LINKE
if (party.selec.switch==8) party.selection <- c(1:2,4:7)   # Without GRUENE
if (party.selec.switch==9) party.selection <- c(1:4,6:7)   # Without AfD
if (party.selec.switch==10) party.selection <- c(1,3:7)    # Without CDU/CSU
if (party.selec.switch==11) party.selection <- c(2:4,6:7)  # Without SPD and DIE LINKE
if (party.selec.switch==12) party.selection <- c(1:6)      # Without BSW (traditional parties only)
```


```{r}
all.abs.votes <- all.abs.votes[,party.selection,,,]
all.treat <- all.treat[,party.selection,,]

```


## Switch for denominator to compute the vote share

  1: eligible citizens (for turnout in analysis, this option has to be selected.)
  2: all voters
  3: all valid votes 

```{r}

vote.denom <- 1

```

### Election results

We first compute the relative vote shares: 

```{r}

  all.rel.votes <-  all.abs.votes/all.abs.total[,rep(vote.denom,length(party.selection)),,,] 
  # If one uses valid votes: rep(3,6)
  # If one uses all voters: rep(2,6)
  # If one uses all eligible citizens: rep(1,6)

  # We delete the past results
  all.rel.votes <- all.rel.votes[,,,,1]

```



### Treatment

```{r}

all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy>0] <- 1

table(all.treat.dummy)



```

## index array

Year index

```{r}
year.index <- all.treat.dummy

year.index[, , , 1] <- 2025
year.index[, , , 2] <- 2021
year.index[, , , 3] <- 2017
year.index[, , , 4] <- 2013
year.index[, , , 5] <- 2009
year.index[, , , 6] <- 2005
year.index[, , , 7] <- 2002
# 
table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy

party.labels <- c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW")[party.selection] # OLD: c("CDU","SPD","AFD","FDP","LIN","GRU")
for (i.party in 1:length(party.selection)){
  party.index[,i.party,,] <- party.labels[i.party]
}

table(party.index)
```

Modification of Wahlkreis index array

```{r}
all.wkr.index <- all.wkr.index[,party.selection,,] # Wenever adjust for the retraced WK boundaries?

```


### Special treatment of Greens (2021)

```{r}
print(paste("Number of non-missings",sum(is.na(all.rel.votes))))
if (party.selec.switch==2) {
  all.rel.votes[party.index=="GRU" & year.index<2017] <- NA
}
if (party.selec.switch==3) {
  all.rel.votes[party.index=="GRU" & year.index==2021] <- NA
}

print(paste("Number of non-missings after making Baerbock as KK",sum(is.na(all.rel.votes))))


```


# Fixed effect models

```{r}
library(lmtest)
library(sandwich)

for (i.smdpr in 1:2){ # Loop over Erst- and Zweitstimme

  all.results <- vector("list",6) # empty list to store the results

  votes.in.array <- all.rel.votes[,,i.smdpr,1:6]
  votes <- c(all.rel.votes[,,i.smdpr,1:6] )
  last.votes <- c(all.rel.votes[,,i.smdpr,2:7] )
  treat <- c(all.treat.dummy[,,i.smdpr,1:6] )
  year  <- as.factor(c(year.index[,,i.smdpr,1:6] ))
  year <- relevel(year,ref="2005") # Why 2005?
  party <- factor(c(party.index[, , i.smdpr, 1:6]),
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW")) # Fixed order and added BSW
  party <- relevel(party, ref = "CDU")
  district <- as.factor(c(all.wkr.index[,,i.smdpr,1:6]))
  cross.treat <- c(all.treat.dummy[,,c(2,1)[i.smdpr],1:6] )

  lm.dat <- as.data.frame(votes)
  lm.dat <- cbind(lm.dat,treat,cross.treat,last.votes,year,party,district)
  
  print(paste("Number of observations: ",nrow(na.omit(lm.dat))))
  
  print("Number of observations per party: ")
  print(table(na.omit(lm.dat)$party))
  
  
  ## Model with year and party fixed effects.
  print(c("Erststimme (H1a)","Zweitstimme  (H1b)")[i.smdpr])
  options(max.print=100)
  print(summary(lm.fe1 <- lm(votes ~ treat + last.votes+
                               party * year + district*party )))

  print(c("Robust standard error"))
  r.se.fe1 <- sqrt(diag(vcovHC(lm.fe1, type = 'HC')))
  print(coef.fe1 <- coeftest(lm.fe1, vcov = vcovHC(lm.fe1, type = 'HC')))

  all.results[[6]] <- lm.fe1
  
  
  ## Model with cross treatment (contamination)

  print(c("Contamination Erststimme (H2b)","Contamination Zweitstimme  (H2a)")[i.smdpr])
  print(summary(lm.fe2 <- lm(votes~ treat + cross.treat + 
                               last.votes + party * year + district * party)))
  
  print(c("Robust standard error"))
  r.se.fe2 <- sqrt(diag(vcovHC(lm.fe2, type = 'HC')))
  print(lm.fe2.robust <- coeftest(lm.fe2, vcov = vcovHC(lm.fe2, type = 'HC')))
  
  
  all.results[[1]] <- lm.fe2
  all.results[[2]] <- lm.fe2.robust
  
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  
  all.results[[5]] <- r.se.fe2
  
  if (i.smdpr==1) all.results.smd <- all.results
  if (i.smdpr==2) all.results.pr <- all.results
}

save(all.results.smd,
     all.results.pr,
 file=paste0("Data/Estimation_Results_FE_RE_PSelect",party.selec.switch,
              "_denom",vote.denom,".RData")
              )

options(max.print=10000)
```

## Some diagnostic checks


Checking residuals

```{r}

results.te <- all.results.smd[[1]] # nur für Erststimme? 

#plot(results.te)


data.te <- all.results.smd[[3]]

data.te <- data.te[!is.na(data.te$votes)&!is.na(data.te$last.votes),]
dim(data.te)


residuals <- residuals(results.te)
predicted <- predict(results.te)

length(residuals)


plot(residuals ~ data.te$treat)
tapply(residuals , as.factor( data.te$treat),mean)



```

### Plotting the actual votes, predicted votes and residuals for treated district

```{r}
par(mfrow=c(1,2)) 

plot(data.te$votes[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted",ylab="Actual")
abline(coef=c(0,1),lty=2)
labels <- data.te[data.te$treat==1,3:5]

text( predicted[data.te$treat==1],data.te$votes[data.te$treat==1] ,
     rownames(data.te)[data.te$treat==1],
     pos=1,cex=0.7)

summary(lm.rsd <- lm(residuals[data.te$treat==1] ~ predicted[data.te$treat==1]))
plot(residuals[data.te$treat==1] ~ predicted[data.te$treat==1],
     xlab="Predicted",ylab="Rersiduals")
abline(lm.rsd)

labels <- data.te[data.te$treat==1,3:5]

text( predicted[data.te$treat==1],residuals[data.te$treat==1] ,
     rownames(data.te)[data.te$treat==1],
     pos=1,cex=0.7)


labels

```


```{r}

wkr.fe <- coefficients(results.te)
wkr.fe <- wkr.fe[grep("district",names(wkr.fe))]

hist(wkr.fe)

wkr.fe[order(wkr.fe)][1:5]

wkr.fe[order(wkr.fe,decreasing=TRUE)][1:5]

```


### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}

for (i.smdpr in 1:2){
  if (i.smdpr==1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]

  }else{
    to.be.tested <- all.results.pr[[1]] # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  

residuals <- residuals(to.be.tested)
residuals <- array(residuals,dim=dim(temp.depv))

simple.district.idx <- all.wkr.index[,1,1,]

# stacking residuals at t and t-1
## loop from BTW 2021 to 2009

current.resid <- past.resid <- residuals[,,1:4]
for (i.btw in 1:4){
  # create the key to match the current and past districts
  key <- match(simple.district.idx[,i.btw],simple.district.idx[,i.btw+1])
  past.resid[,,i.btw] <- residuals[key,,i.btw+1]
}

print(c("Erststimme","Zweitstimme")[i.smdpr])
print(
   summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
)
plot(c(current.resid) ~ c(past.resid),
     ylab="t",xlab="t-1")
abline(lm.out)

}

```


### Exogeneity

```{r}
# Wooldridge's test

options(max.print=100)
# District candidacy: Treatment for SMD
 
dep.var <- c(all.rel.votes[,,1,2:5])
lagged.dv <- c(all.rel.votes[,,1,3:6])
treat <- c(all.rel.votes[,,1,2:5])
treat.t1 <- c(all.rel.votes[,,1,1:4])
cross.treat <- c(all.rel.votes[,,2,2:5])
cross.treat.t1 <- c(all.rel.votes[,,2,1:4])

party.fe <- factor(c(party.index[,,1,2:5]),
                   levels=c("CDU","SPD","AFD","FDP","LIN","GRU"))
wkr.fe <- as.factor(c(all.wkr.index[,,1,2:5]))

year.fe  <- as.factor(c(year.index[,,1,1:4] ))

summary(lm.out <- lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))

 print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))

all.results.smd <- vector("list",3)
all.results.smd[[1]] <- lm.out
all.results.smd[[2]] <- lm.out.robust
all.results.smd[[3]] <- c("empty")
all.results.smd[[4]] <- c("empty")
all.results.smd[[5]] <- r.se


# PR list candidacy: Treatment for PR
 
all.results <- vector("list",3)

dep.var <- c(all.rel.votes[,,2,2:5])
lagged.dv <- c(all.rel.votes[,,2,3:6])
treat <- c(all.rel.votes[,,2,2:5])
treat.t1 <- c(all.rel.votes[,,2,1:4])
cross.treat <- c(all.rel.votes[,,1,2:5])
cross.treat.t1 <- c(all.rel.votes[,,1,1:4])

party.fe <- factor(c(party.index[,,2,2:5]),
                   levels=c("CDU","SPD","AFD","FDP","LIN","GRU"))
wkr.fe <- as.factor(c(all.wkr.index[,,2,2:5]))

year.fe  <- as.factor(c(year.index[,,2,1:4] ))

summary(lm.out <- lm(dep.var ~ lagged.dv +
             treat + treat.t1+
             cross.treat + cross.treat.t1+
             year.fe*party.fe + party.fe*wkr.fe))

print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))

  
all.results.pr <- vector("list",3)
all.results.pr[[1]] <- lm.out
all.results.pr[[2]] <- lm.out.robust
all.results.pr[[3]] <- c("empty")
all.results.pr[[4]] <- c("empty")
all.results.pr[[5]] <- r.se


save(all.results.smd,
     all.results.pr,
 file=(paste0("Data/Exogeneity_Results_FE_PSelect",party.selec.switch,
              "_denom",vote.denom,
                ".RData")))

options(max.print=10000)

```



