---
title: "First Differences"
author: "Anna Werner & Susumu Shikano"
date: "`r Sys.Date()`"
output: html_document
---

```{r , include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stargazer)
library(MASS)
```


## Load data

```{r}
load("Data/analysis_dat.RData")
```

```{r}
# save ouputs
save.outputs <- 1
```

## Switch for party selection and spill over
 
```{r}
party.selec.switch <- 1
spill.over.switch <- 0 # 1: with spill over
turnout.switch <- 0 # 1: turnout analysis
```

See below what value corresponds to what party selection.

```{r}
# OLD !!!

#if (party.selec.switch==1) party.selection <- c(1:6)      # S A F L G 
#if (party.selec.switch==2) party.selection <- c(1:2,6)      # nur C und S + G (2021)
#if (party.selec.switch==3) party.selection <- c(3:6)      # ohne C und S und G (2021)
#if (party.selec.switch==4) party.selection <- c(2:6)      # ohne C
#if (party.selec.switch==5) party.selection <- c(1,3:6)    # ohne S
#if (party.selec.switch==6) party.selection <- c(1:3,5:6) # ohne F
#if (party.selec.switch==7) party.selection <- c(1:4,6)    # ohne Linke
#if (party.selec.switch==8) party.selection <- c(1:5)     # ohne G
#if (party.selec.switch==9) party.selection <- c(1:2,4:6)  # ohne AfD
#if (party.selec.switch==10) party.selection <- c(2,4:6)    # ohne C und A
#if (party.selec.switch==11) party.selection <- c(3:4,6)    # ohne C, S und L


## Do not select only one party. This won't work!
## Party positions: 1=SPD, 2=CDU/CSU, 3=GRUENE, 4=FDP, 5=AfD, 6=DIE LINKE, 7=BSW

if (party.selec.switch==1) party.selection <- c(1:7)       # All parties: S C G F A L B
if (party.selec.switch==2) party.selection <- c(1:2,3)     # SPD, CDU/CSU, GRUENE
if (party.selec.switch==3) party.selection <- c(4:7)       # FDP, AfD, DIE LINKE, BSW (without SPD, CDU/CSU, GRUENE)
if (party.selec.switch==4) party.selection <- c(1,3:7)     # Without CDU/CSU
if (party.selec.switch==5) party.selection <- c(2:7)       # Without SPD
if (party.selec.switch==6) party.selection <- c(1:3,5:7)   # Without FDP
if (party.selec.switch==7) party.selection <- c(1:5,7)     # Without DIE LINKE
if (party.selec.switch==8) party.selection <- c(1:2,4:7)   # Without GRUENE
if (party.selec.switch==9) party.selection <- c(1:4,6:7)   # Without AfD
if (party.selec.switch==10) party.selection <- c(1,3:7)    # Without CDU/CSU
if (party.selec.switch==11) party.selection <- c(2:4,6:7)  # Without SPD and DIE LINKE
if (party.selec.switch==12) party.selection <- c(1:6)      # Without BSW (traditional parties only)
```

Reading adjacent matrix

```{r}
if (spill.over.switch>0){
  load("Data/btw_all_shape.RData")
  all.nb.mat <- all.nb.mat[3:7] # <- update with new dat???
}

```



## Switch for denominator to compute the vote share

  1: eligible citizens (for turnout in analysis, this option has to be selected.)
  2: all voters
  3: all valid votes 

```{r}

vote.denom <- 1
if (turnout.switch==1) vote.denom <- 1 

```


all.abs.votes has 5 dimensions:
* **Constituencies** (1 … 299)
* **Parties** (1 … 7): S C G F A L B
* **First (FPTP) or second (PR) vote** (1, 2)
* **Year** (1 … 7): 2025, 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

all.treat has 4 dimensions:
* **Constituencies** (1 … 299)
* **Parties** (1 … 7): S C G F A L B
* **First or second vote** (1, 2)
* **Year** (1 … 7): 2025, 2021, 2017, 2013, 2009, 2005, 2002

```{r}
# UPDATE for new year
if (turnout.switch==0){
  all.abs.votes <- all.abs.votes[,party.selection,,,]
  all.treat     <- all.treat[,party.selection,,]
  all.rel.votes <-  all.abs.votes/
  all.abs.total[,rep(vote.denom,length(party.selection)),,,,drop=F] 
}else{
  all.abs.votes <- all.abs.total[,2,,,,drop=F] 
  all.treat     <- all.treat[,party.selection,,]
  temp <- apply(all.treat,c(1,3,4),max)
  all.treat <- array(temp,dim=c(299,1,2,6))    
  all.rel.votes <-  all.abs.votes/all.abs.total[,rep(vote.denom,1),,,,drop=F] 
}

```


## Creating first difference

### Election results

We create the first differences:


```{r}
# First differences: current results minus previous results
# Dimensions: [constituencies, parties, vote_type, years, current_vs_previous]
# We can compute FD for years 1-6 (2025, 2021, 2017, 2013, 2009, 2005)
# No FD for year 7 (2002) because there's no "previous" data

all.rel.votes.fd <- all.rel.votes[, , , 1:6, 1, drop=FALSE] - 
                    all.rel.votes[, , , 1:6, 2, drop=FALSE]

# Note that there will be no FD for 2002, therefore we have n(year) = 6
all.rel.votes.fd <- array(all.rel.votes.fd,
                          dim = dim(all.rel.votes.fd)[1:4])

hist(all.rel.votes.fd)

# Check dimensions
dim(all.rel.votes.fd)  # Should be: 299, n_parties, 2, 6
```

### Treatment

Treatment leading candidate or not

```{r}
# Make all values > 0 to 1 (convert treatment levels to binary indicator)
all.treat.dummy <- all.treat
all.treat.dummy[all.treat.dummy > 0] <- 1

# Incumbent treatment: only values 3 or 4 indicate incumbent
all.treat.incumb <- ifelse(all.treat == 3 | all.treat == 4, 1, 0)

# Check distributions
table(all.treat.dummy)
apply(all.treat.dummy, 3, table)
table(all.treat.incumb)
apply(all.treat.incumb, 3, table)

# Create first difference arrays for treatments
# Now 6 time periods (2025-2021, 2021-2017, 2017-2013, 2013-2009, 2009-2005, 2005-2002)
if (turnout.switch == 0){
  all.treat.dummy.fd <- 
  all.treat.incumb.fd <- 
      array(NA, dim = c(299, length(party.selection), 2, 6))  # Changed from 5 to 6
} else {
  all.treat.dummy.fd <- 
  all.treat.incumb.fd <- 
      array(NA, dim = c(299, 1, 2, 6))  # Changed from 5 to 6
}

# Compute first differences for treatments
for (i.time in 1:6){  # Changed from 1:5 to 1:6
  all.treat.dummy.fd[, , , i.time] <-  
    all.treat.dummy[, , , i.time] - all.treat.dummy[, , , i.time + 1] 
  all.treat.incumb.fd[, , , i.time] <-  
    all.treat.incumb[, , , i.time] - all.treat.incumb[, , , i.time + 1] 
}

# Check first difference distributions
table(all.treat.dummy.fd)
apply(all.treat.dummy.fd, 3, table)
table(all.treat.incumb.fd)
apply(all.treat.incumb.fd, 3, table)
```

The resulting all.treat.dummy.fd has:
- 299 districts
- 7 parties
- 2 tiers
- 6 timepoints


Incorporating spill-over treatment

```{r}
if (spill.over.switch > 0){
  for (i.tier in 1:2){
    for (i.btw in 1:6){  # Changed from 1:5 to 1:6
      temp.treat <- all.treat.dummy.fd[, , i.tier, i.btw]
      temp.mat <- as.matrix(all.nb.mat[[7 - i.btw]])  # Changed from 6-i.btw to 7-i.btw
      diag(temp.mat) <- 1  # replace the diagonal elements with 1
      spill.treat <- temp.mat %*% temp.treat
      spill.treat[spill.treat > 0] <- 1
      spill.treat[spill.treat < 0] <- -1
      all.treat.dummy.fd[, , i.tier, i.btw] <- spill.treat
    }
  }
}

apply(all.treat.dummy.fd, 3, table)
```




## index array

Year index

```{r}
year.index <- all.treat.dummy.fd
year.index[, , , 1] <- 2025
year.index[, , , 2] <- 2021
year.index[, , , 3] <- 2017
year.index[, , , 4] <- 2013
year.index[, , , 5] <- 2009
year.index[, , , 6] <- 2005
table(year.index)

```

Party index

```{r}
party.index <- all.treat.dummy.fd

if (turnout.switch == 0){
  # Party order in arrays: SPD, CDU/CSU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party.labels <- c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW")[party.selection]
  
  for (i.party in 1:length(party.selection)){
    party.index[, i.party, , ] <- party.labels[i.party]
  }
} else {
  party.labels <- "Turnout"
  party.index[, 1, , ] <- party.labels
}

table(party.index)
```

### Special treatment of Greens (2021)

```{r}
if (party.selec.switch==2) {
  all.rel.votes.fd[party.index=="GRU" & year.index!=2021] <- NA
}
if (party.selec.switch==3) {
  all.rel.votes.fd[party.index=="GRU" & year.index==2021] <- NA
}

```




## First difference models

```{r}
library(sandwich)
library(lmtest)

# Optionally: previousl we used HC SEs. But those may not adequately deal with the autocorrelation.
# Why not clustered SEs by constituency? Uncomment that option, and comment out the HC approach, if wanted.

for (i.smdpr in 1:2){  # Loop over Erst- and Zweitstimme
  all.results <- vector("list", 5)  # empty list to store the results
  votes.in.array <- all.rel.votes.fd[, , i.smdpr, ]
  votes <- c(all.rel.votes.fd[, , i.smdpr, ])
  treat <- c(all.treat.dummy.fd[, , i.smdpr, ])
  treat.incumb <- c(all.treat.incumb.fd[, , i.smdpr, ])
  year  <- as.factor(c(year.index[, , i.smdpr, ]))
  year <- relevel(year, ref = "2009")  # Changed from "2005" to "2009" (middle year)
  
  # Party order: SPD, CDU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party <- factor(c(party.index[, , i.smdpr, ]),
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW", "Turnout"))
  party <- relevel(party, ref = "AFD")
  
  cross.treat <- c(all.treat.dummy.fd[, , c(2,1)[i.smdpr], ])
  cross.treat.incumb <- c(all.treat.incumb.fd[, , c(2,1)[i.smdpr], ])
  
  # Each of the 299 constituencies appears multiple times (once per party per time period)
  constituency <- rep(1:299, times = length(party.selection) * 6) # new
  
  lm.dat <- cbind(votes, treat, cross.treat, treat.incumb, cross.treat.incumb)
  lm.dat <- as.data.frame(lm.dat)
  lm.dat$year <- year
  lm.dat$party <- party
  lm.dat$constituency <- constituency  # NEW: Add constituency ID
  lm.dat <- na.omit(lm.dat)
  
  ## Model 
  print(c("DV: Erststimme", "DV: Zweitstimme")[i.smdpr])
  if (turnout.switch == 0){
    print(summary(lm.out <- lm(votes ~ treat +
                                  #treat.incumb + # control for Incumbency
                                  cross.treat +
                                  #cross.treat.incumb +
                                  year + year:party 
                                  , data = lm.dat))) 
  } else {
    print(summary(lm.out <- lm(votes ~ treat +
                                  #treat.incumb + # control for Incumbency
                                  cross.treat +
                                  #cross.treat.incumb +
                                  year 
                                  , data = lm.dat))) 
  }
  
  # HC option
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  
  # OPTION: Cluster-robust standard errors
  #print("\n Cluster-Robust SEs (by Constituency)\n")
  #print("Accounts for serial correlation and heteroskedasticity within constituencies\n")
  #vcov_cluster <- vcovCL(lm.out, cluster = lm.dat$constituency, type = "HC1")
  #lm.out.cluster <- coeftest(lm.out, vcov = vcov_cluster)
  #print(lm.out.cluster)
  #r.se.cluster <- sqrt(diag(vcov_cluster))
  
  all.results[[1]] <- lm.out
  all.results[[2]] <- lm.out.robust
  #all.results[[2]] <- lm.out.cluster
  all.results[[3]] <- lm.dat
  all.results[[4]] <- votes.in.array
  all.results[[5]] <- r.se
  #all.results[[5]] <- r.se.cluster
  
  if (i.smdpr == 1) all.results.smd <- all.results
  if (i.smdpr == 2) all.results.pr <- all.results
}
```


```{r}
if (save.outputs == 1) {
  save(
    all.results.smd,
    all.results.pr,
    file = paste0(
      "Data/Estimation_Results_25/First_Differencing/Estimation_Results_PSelect",
      party.selec.switch,
      "_denom", vote.denom,
      c("", "_spillover")[spill.over.switch + 1],
      c("", "_turnout")[turnout.switch + 1],
      "_", format(Sys.Date(), "%Y-%m-%d"),
      ".RData"
    )
  )
}
```


```{r}
# Residuals vs Fitted
plot(lm.out, which = 1)  # Should show random scatter if homoskedastic

# Scale-Location plot
plot(lm.out, which = 3)  # Should show horizontal line if homoskedastic

# Studentized residuals
plot(studres(lm.out) ~ fitted(lm.out),
     xlab = "Fitted Values",
     ylab = "Studentized Residuals",
     main = "Residuals vs Fitted")
abline(h = 0, lty = 2)
```

```{r}
## Diagnostic Tests
print("Diagnostic Tests")


for (i.smdpr in 1:2){
  if (i.smdpr == 1){
    lm.out <- all.results.smd[[1]]
    temp.dat <- all.results.smd[[3]]
  } else {
    lm.out <- all.results.pr[[1]]
    temp.dat <- all.results.pr[[3]]
  }
  
  print("")
  print(c("ERSTSTIMME", "ZWEITSTIMME")[i.smdpr])
  print("")
  
  # Breusch-Pagan Test
  print("Breusch-Pagan Test (Heteroskedasticity):")
  bp_test <- bptest(lm.out)
  print(bp_test)
  
  # Visual diagnostic
  par(mfrow = c(2, 2))
  plot(lm.out)
  par(mfrow = c(1, 1))
  
  # Shapiro-Wilk (normality)
  print("Shapiro-Wilk Test (Normality of Residuals):")
  shapiro.test(sample(residuals(lm.out), 5000))  # if n > 5000
}
```



```{r}
if (turnout.switch==1|spill.over.switch==1) knitr::knit_exit()
```

### Diagnostics I: Serial correlated errors

We regress u_t on u_t-1. If the effect is significant, there are serially correlated errors.

Note: We first have to assign individual districts at t to those at t-1.

```{r}
for (i.smdpr in 1:2){
  if (i.smdpr == 1){
    to.be.tested <- all.results.smd[[1]]  # Diagnostics for the model Erststimme
    temp.dat <- all.results.smd[[3]]
    temp.depv <- all.results.smd[[4]]
  } else {
    to.be.tested <- all.results.pr[[1]]  # Diagnostics for the model Zweitstimme
    temp.dat <- all.results.pr[[3]]
    temp.depv <- all.results.pr[[4]]
  }
  
  residuals <- residuals(to.be.tested)
  if (party.selec.switch != 0){
    residuals <- array(residuals, dim = dim(temp.depv))
  } else {
    residuals <- array(residuals,
                       dim = c(dim(temp.depv)[1], 1, dim(temp.depv)[2]))
  }
  
  simple.district.idx <- all.wkr.index[, 1, 1, ]
  
  # Stacking residuals at t and t-1
  ## Loop from BTW 2025 to 2005 (6 time periods, 5 pairs)
  current.resid <- past.resid <- residuals[, , 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  
  for (i.btw in 1:5){  # Changed from 1:4 to 1:5
    # Create the key to match the current and past districts
    key <- match(simple.district.idx[, i.btw], simple.district.idx[, i.btw + 1])
    past.resid[, , i.btw] <- residuals[key, , i.btw + 1]
  }
  
  print(c("Erststimme", "Zweitstimme")[i.smdpr])
  print(
    summary(lm.out <- lm(c(current.resid) ~ c(past.resid)))
  )
  
  plot(c(current.resid) ~ c(past.resid),
       ylab = "t", xlab = "t-1",
       main = paste(c("Erststimme", "Zweitstimme")[i.smdpr], 
                    "- Serial Correlation Test"))
  abline(lm.out)
}
```
Consider: Does the negative serial correlation make theoretical sense? It might indicate:
- Mean reversion in vote shares
- Voters "correcting" previous swings
- Backlash effects


### Diagnostics II: Exogeneity of the treatment

We are testing by using Wooldridge's test. That is, we include the treatment variable at t+1. For this purpose, we lose the most recent BTW 2021.

Also here, we have to take care about the redistricting.

```{r}
for (i.smdpr in 1:2){  # Loop over Erst- and Zweitstimme

  all.results.t1 <- vector("list", 5)  # empty list to store the results (changed from 3 to 5)

  votes.in.array <- all.rel.votes.fd[, , i.smdpr, 2:6]  # Changed from 2:5 to 2:6
  votes <- c(votes.in.array)
  treat <- c(all.treat.dummy.fd[, , i.smdpr, 2:6])  # Changed from 2:5 to 2:6
  treat.incumb <- c(all.treat.incumb.fd[, , i.smdpr, 2:6])  # Changed from 2:5 to 2:6
  year  <- as.factor(c(year.index[, , i.smdpr, 2:6]))  # Changed from 2:5 to 2:6
  year <- relevel(year, ref = "2009")  # Changed from "2005" to "2009"
  
  # Party order: SPD, CDU, GRUENE, FDP, AfD, DIE LINKE, BSW
  party <- factor(c(party.index[, , i.smdpr, 2:6]),  # Changed from 2:5 to 2:6
                  levels = c("SPD", "CDU", "GRU", "FDP", "AFD", "LIN", "BSW"))
  party <- relevel(party, ref = "AFD")
  
  cross.treat <- c(all.treat.dummy.fd[, , c(2,1)[i.smdpr], 2:6])  # Changed from 2:5 to 2:6
  cross.treat.incumb <- c(all.treat.incumb.fd[, , c(2,1)[i.smdpr], 2:6])  # Changed from 2:5 to 2:6
  
  # Treatment at t+1
  treat.t1 <- all.treat.dummy.fd[, , i.smdpr, 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  treat.t1 <- array(treat.t1, dim = dim(treat.t1)[c(1,2,4)])
  treat.incumb.t1 <- all.treat.incumb.fd[, , i.smdpr, 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  treat.incumb.t1 <- array(treat.incumb.t1, dim = dim(treat.incumb.t1)[c(1,2,4)])
  cross.treat.t1 <- all.treat.dummy.fd[, , c(2,1)[i.smdpr], 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  cross.treat.t1 <- array(cross.treat.t1, dim = dim(cross.treat.t1)[c(1,2,4)])
  cross.treat.incumb.t1 <- all.treat.incumb.fd[, , c(2,1)[i.smdpr], 1:5, drop = FALSE]  # Changed from 1:4 to 1:5
  cross.treat.incumb.t1 <- array(cross.treat.incumb.t1, dim = dim(cross.treat.incumb.t1)[c(1,2,4)])
  
  # We assign the correct district
  simple.district.idx <- all.wkr.index[, 1, 1, ]
  for (i.btw in 2:6){  # Changed from 2:5 to 2:6 (loop over btw2021 to btw2005)
    key <- match(simple.district.idx[, i.btw], simple.district.idx[, i.btw - 1])
    treat.t1[, , i.btw - 1] <- treat.t1[key, , i.btw - 1]
    treat.incumb.t1[, , i.btw - 1] <- treat.incumb.t1[key, , i.btw - 1]
    cross.treat.t1[, , i.btw - 1] <- cross.treat.t1[key, , i.btw - 1]
    cross.treat.incumb.t1[, , i.btw - 1] <- cross.treat.incumb.t1[key, , i.btw - 1]
  }
  treat.t1 <- c(treat.t1)
  cross.treat.t1 <- c(cross.treat.t1)
  treat.incumb.t1 <- c(treat.incumb.t1)
  cross.treat.incumb.t1 <- c(cross.treat.incumb.t1)

  lm.dat <- cbind(votes, treat, year, party,
                  cross.treat, treat.incumb, cross.treat.incumb,
                  treat.t1, cross.treat.t1, treat.incumb.t1, cross.treat.incumb.t1)
  lm.dat <- as.data.frame(lm.dat)
  
  ## Model with cross treatment (contamination)

  print(c("DV: Erststimme", "DV: Zweitstimme")[i.smdpr])
  if (party.selec.switch != 0){
    print(summary(lm.out <- lm(votes ~ treat + treat.t1 +
                                  #treat.incumb + 
                                  #treat.incumb.t1 + # control for Incumbency
                                  cross.treat + cross.treat.t1 +
                                  #cross.treat.incumb + cross.treat.incumb.t1 +
                                  as.factor(year) + as.factor(year):as.factor(party),
                               data = lm.dat)))
  } else {
    print(summary(lm.out <- lm(votes ~ treat + treat.t1 +
                                  #treat.incumb + 
                                  #treat.incumb.t1 + # control for Incumbency
                                  cross.treat + cross.treat.t1 +
                                  #cross.treat.incumb + cross.treat.incumb.t1 +
                                  as.factor(year),
                               data = lm.dat)))
  }
  
  print(c("Robust standard error"))
  r.se <- sqrt(diag(vcovHC(lm.out, type = 'HC')))
  print(lm.out.robust <- coeftest(lm.out, vcov = vcovHC(lm.out, type = 'HC')))
  
  print(c("Treatments Erststimmen", "Treatments Zweitstimmen")[i.smdpr])     
  print(table(treat))
  print(table(treat.t1))
  print(table(cross.treat))
  print(table(cross.treat.t1))

  all.results.t1[[1]] <- lm.out
  all.results.t1[[2]] <- lm.out.robust
  all.results.t1[[3]] <- lm.dat
  all.results.t1[[4]] <- votes.in.array
  all.results.t1[[5]] <- r.se
  
  if (i.smdpr == 1) all.results.t1.smd <- all.results.t1
  if (i.smdpr == 2) all.results.t1.pr <- all.results.t1
}

save(all.results.t1.smd,
     all.results.t1.pr,
     file = paste0("Data/Estimation_Results_25/Exogeneity/Exogeneity_Results_PSelect", 
                   party.selec.switch,
                   "_denom", vote.denom,
                   c("", "_spillover")[spill.over.switch + 1],
                   c("", "_turnout")[turnout.switch + 1],
                   "_", format(Sys.Date(), "%Y%m%d"),
                   ".RData"))

```
