---
title: "Preparing the pooled datasets"
author: "Anna Werner"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## load packages ##
pacman::p_load(readr,foreign)
```


## Dataset

This document contains the R code used to combine the individual CSV files from the Federal Returning Officer, which include the results of in-person and postal voting by constituency for the federal elections from 2002 to 2021, into a single, consolidated dataset. The corresponding CSV files are publicly available on the website of the Federal Returning Officer (Tables: Final results by constituency for all federal elections, [https://www.bundeswahlleiter.de/bundestagswahlen/2021/ergebnisse.html](https://www.bundeswahlleiter.de/bundestagswahlen/2021/ergebnisse.html), as of December 13, 2022).

The final results by federal state are also provided. These were accessed on [date to be inserted].

## Individual datasets by election year
```{r Data25}

### BTW25 ###

# read in results
btw25_raw <- read.csv("Data/btw_kerg/btw2025_kerg_00413.csv",
                      skip=8, sep=";", header = FALSE, encoding="UTF-8")

# Note: additional var "Gewählt"

# read in first row of labels
main.labels25 <- read.csv("Data/btw_kerg/btw2025_kerg_00413.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="UTF-8")

# read in second row of labels
first.second.votes25 <- read.csv("Data/btw_kerg/btw2025_kerg_00413.csv",
                               skip=6, nrows=1, sep=";", header = FALSE)
first.second.votes25[1:4] <- "aaa"

# delete previous results and state-level results
result2525 <- btw25_raw[,!is.na(first.second.votes25)]
result2525 <- result2525[result2525[,3]!=99 & !is.na(result2525[,3]),]

# keep only previous results and delete state-level results
result2521 <- btw25_raw[,is.na(first.second.votes25)|first.second.votes25=="aaa"]
result2521 <- result2521[result2521[,3]!=99 & !is.na(result2521[,3]),]
result2521 <- result2521[,1:ncol(result2525)]

# make labels
main.labels25 <- main.labels25[!is.na(first.second.votes25)]
main.labels25[seq(6, length(main.labels25), by=2)] <-  # one more initial col in 25
  main.labels25[seq(5, length(main.labels25)-1, by=2)] # adding the label for SMD and PR votes

main.labels25 <- paste(main.labels25, first.second.votes25[!is.na(first.second.votes25)])
main.labels25 <- gsub("aaa", "", main.labels25)

main.labels25 <- gsub("ä", "ae", main.labels25)
main.labels25 <- gsub("Ä", "AE", main.labels25)
main.labels25 <- gsub("ö", "oe", main.labels25)
main.labels25 <- gsub("Ö", "OE", main.labels25)
main.labels25 <- gsub("ü", "ue", main.labels25)
main.labels25 <- gsub("Ü", "UE", main.labels25)

colnames(result2525) <- colnames(result2521) <- main.labels25

# paste BSW data behind the major parties
bsw_idx <- grep("Sahra Wagenknecht", names(result2525)) # NOTE: BSW is in position [65][66]
new_order <- c(1:26, bsw_idx, setdiff(seq_along(result2525), c(1:26, bsw_idx)))

result2525 <- result2525[, new_order]
result2521 <- result2521[, names(result2525)]  # align by name



# select columns
result2525 <- result2525[,1:28]  
result2521 <- result2521[,1:28]

# fuse CDU and CSU into CDU/CSU
union2525 <- result2525[,grep("Christlich Demokratische",names(result2525))]
csu2525 <- result2525[,grep("Christlich-Soziale",names(result2525))]
union2525[is.na(union2525[,1]),] <- csu2525[is.na(union2525[,1]),]
result2525[,grep("Christlich Demokratische",names(result2525))] <- union2525

result2525 <- result2525[,grep("Christlich-Soziale",names(result2525),invert=T)]


union2521 <- result2521[,grep("Christlich Demokratische",names(result2521))]
csu2521 <- result2521[,grep("Christlich-Soziale",names(result2521))]
union2521[is.na(union2521[,1]),] <- csu2521[is.na(union2521[,1]),]
result2521[,grep("Christlich Demokratische",names(result2521))] <- union2521

result2521 <- result2521[,grep("Christlich-Soziale",names(result2521),invert=T)]


# drop additional column "Gewählt"
result2525 <- result2525[,-4]
result2521 <- result2521[,-4]


# create variable for year of data
result2525$Jahr <- 2025
result2521$Jahr <- 2025

# rename variables
colnames(result2525)[3] <- "Land"
colnames(result2525)[8:11] <- c("Ungueltige Erststimmen","Ungueltige Zweitstimmen", "Gueltige Erststimmen", "Gueltige Zweitstimmen")
colnames(result2525)[12:25] <- c("SPD Erststimmen","SPD Zweitstimmen","CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen", "BSW Erststimmen", "BSW Zweitstimmen")

colnames(result2521)[3] <- "Land"
colnames(result2521)[8:11] <- c("Ungueltige Erststimmen","Ungueltige Zweitstimmen", "Gueltige Erststimmen", "Gueltige Zweitstimmen")
colnames(result2521)[12:25] <- c("SPD Erststimmen","SPD Zweitstimmen","CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen", "BSW Erststimmen", "BSW Zweitstimmen")



# sort based on the WKR number
result2525 <- result2525[order(result2525$`Nr `),]
result2521 <- result2521[order(result2521$`Nr `),]


```

```{r Data21}

### BTW21 ###

# read in results
btw21_raw <- read.csv("Data/btw_kerg/btw2021_kerg.csv",
                      skip=5, sep=";", header = FALSE, encoding="UTF-8")

# read in first row of labels
main.labels21 <- read.csv("Data/btw_kerg/btw2021_kerg.csv",
                        skip=2, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="UTF-8")

# read in second row of labels
first.second.votes21 <- read.csv("Data/btw_kerg/btw2021_kerg.csv",
                               skip=3, nrows=1, sep=";", header = FALSE)
first.second.votes21[1:3] <- "aaa"

# delete previous results and state-level results
result2121 <- btw21_raw[,!is.na(first.second.votes21)]
result2121 <- result2121[result2121[,3]!=99 & !is.na(result2121[,3]),]

# keep only previous results and delete stat-level results
result2117 <- btw21_raw[,is.na(first.second.votes21)|first.second.votes21=="aaa"]
result2117 <- result2117[result2117[,3]!=99 & !is.na(result2117[,3]),]
result2117 <- result2117[,1:ncol(result2121)]

# make labels
main.labels21 <- main.labels21[!is.na(first.second.votes21)]
main.labels21[seq(5, length(main.labels21), by=2)] <- 
  main.labels21[seq(4, length(main.labels21)-1, by=2)] # adding the label for SMD and PR votes

main.labels21 <- paste(main.labels21, first.second.votes21[!is.na(first.second.votes21)])
main.labels21 <- gsub("aaa", "", main.labels21)

main.labels21 <- gsub("ä", "ae", main.labels21)
main.labels21 <- gsub("Ä", "AE", main.labels21)
main.labels21 <- gsub("ö", "oe", main.labels21)
main.labels21 <- gsub("Ö", "OE", main.labels21)
main.labels21 <- gsub("ü", "ue", main.labels21)
main.labels21 <- gsub("Ü", "UE", main.labels21)

colnames(result2121) <- colnames(result2117) <- main.labels21

# select columns
result2121 <- result2121[,1:25]
result2117 <- result2117[,1:25]

# fuse CDU and CSU into CDU/CSU
union2121 <- result2121[,grep("Christlich Demokratische",names(result2121))]
csu2121 <- result2121[,grep("Christlich-Soziale",names(result2121))]
union2121[is.na(union2121[,1]),] <- csu2121[is.na(union2121[,1]),]
result2121[,grep("Christlich Demokratische",names(result2121))] <- union2121

result2121 <- result2121[,grep("Christlich-Soziale",names(result2121),invert=T)]


union2117 <- result2117[,grep("Christlich Demokratische",names(result2117))]
csu2117 <- result2117[,grep("Christlich-Soziale",names(result2117))]
union2117[is.na(union2117[,1]),] <- csu2117[is.na(union2117[,1]),]
result2117[,grep("Christlich Demokratische",names(result2117))] <- union2117

result2117 <- result2117[,grep("Christlich-Soziale",names(result2117),invert=T)]

# create variable for year of data
result2121$Jahr <- 2021
result2117$Jahr <- 2021

# rename variables
colnames(result2121)[3] <- "Land"
colnames(result2121)[8:11] <- c("Ungueltige Erststimmen","Ungueltige Zweitstimmen", "Gueltige Erststimmen", "Gueltige Zweitstimmen")
colnames(result2121)[12:23] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","SPD Erststimmen","SPD Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen", "GRUENE Erststimmen", "GRUENE Zweitstimmen")

colnames(result2117)[3] <- "Land"
colnames(result2117)[8:11] <- c("Ungueltige Erststimmen","Ungueltige Zweitstimmen", "Gueltige Erststimmen", "Gueltige Zweitstimmen")
colnames(result2117)[12:23] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","SPD Erststimmen","SPD Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen", "GRUENE Erststimmen", "GRUENE Zweitstimmen")



# sort based on the WKR number
result2121 <- result2121[order(result2121$`Nr `),]
result2117 <- result2117[order(result2117$`Nr `),]

```

```{r Data17}
### BTW17 ###

# read in results
btw17_raw <- read.csv("Data/btw_kerg/btw2017_kerg.csv",
                      skip=8, sep=";", header = FALSE, encoding="UTF-8")

# read in first row of labels
main.labels17 <- read.csv("Data/btw_kerg/btw2017_kerg.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="UTF-8")

# read in second row of labels
first.second.votes17 <- read.csv("Data/btw_kerg/btw2017_kerg.csv",
                               skip=6, nrows=1 ,sep=";", header = FALSE)
first.second.votes17[1:3] <- "aaa"

# delete previous results and state-level results
result1717 <- btw17_raw[,!is.na(first.second.votes17)]
result1717 <- result1717[result1717[,3]!=99 & !is.na(result1717[,3]),]

# keep only previous results and delete stat-level results
result1713 <- btw17_raw[,is.na(first.second.votes17)|first.second.votes17=="aaa"]
result1713 <- result1713[result1713[,3]!=99 & !is.na(result1713[,3]),]
result1713 <- result1713[,1:ncol(result1717)]

# make labels
main.labels17 <- main.labels17[!is.na(first.second.votes17)]
main.labels17[seq(5, length(main.labels17), by=2)] <- 
  main.labels17[seq(4, length(main.labels17)-1, by=2)] # adding the label for SMD and PR votes

main.labels17 <- paste(main.labels17, first.second.votes17[!is.na(first.second.votes17)])
main.labels17 <- gsub("aaa", "", main.labels17)

main.labels17 <- gsub("ä", "ae", main.labels17)
main.labels17 <- gsub("Ä", "AE", main.labels17)
main.labels17 <- gsub("ö", "oe", main.labels17)
main.labels17 <- gsub("Ö", "OE", main.labels17)
main.labels17 <- gsub("ü", "ue", main.labels17)
main.labels17 <- gsub("Ü", "UE", main.labels17)

colnames(result1717) <- colnames(result1713) <- main.labels17

# select columns
result1717 <- result1717[,1:25]
result1713 <- result1713[,1:25]

# fuse CDU and CSU into CDU/CSU
union1717 <- result1717[,grep("Christlich Demokratische",names(result1717))]
csu1717 <- result1717[,grep("Christlich-Soziale",names(result1717))]
union1717[is.na(union1717[,1]),] <- csu1717[is.na(union1717[,1]),]
result1717[,grep("Christlich Demokratische",names(result1717))] <- union1717

result1717 <- result1717[,grep("Christlich-Soziale",names(result1717),invert=T)]


union <- result1713[,grep("Christlich Demokratische",names(result1713))]
csu <- result1713[,grep("Christlich-Soziale",names(result1713))]
union[is.na(union[,1]),] <- csu[is.na(union[,1]),]
result1713[,grep("Christlich Demokratische",names(result1713))] <- union

result1713 <- result1713[,grep("Christlich-Soziale",names(result1713),invert=T)]

# create variable for year of data
result1717$Jahr <- 2017
result1713$Jahr <- 2017

# rename variables
colnames(result1717)[3] <- "Land"
colnames(result1717)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result1717)[12:23] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","SPD Erststimmen","SPD Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen")

colnames(result1713)[3] <- "Land"
colnames(result1713)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result1713)[12:23] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","SPD Erststimmen","SPD Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","AfD Erststimmen","AfD Zweitstimmen")


# sort based on the WKR number
result1717 <- result1717[order(result1717$`Nr `),]
result1713 <- result1713[order(result1713$`Nr `),]


```

```{r Data13}
### BTW13 ###

# read in results # AFD eiinfügen
btw13_raw <- read.csv("Data/btw_kerg/btw2013_kerg.csv",
                      skip=8, sep=";", header = FALSE, encoding="latin1")

# read in first row of labels
main.labels13 <- read.csv("Data/btw_kerg/btw2013_kerg.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="latin1")

# read in second row of labels
first.second.votes13 <- read.csv("Data/btw_kerg/btw2013_kerg.csv",
                               skip=6, nrows=1, sep=";", header = FALSE)
first.second.votes13[1:3] <- "aaa"

# delete previous results and state-level results
result1313 <- btw13_raw[,!is.na(first.second.votes13)]
result1313 <- result1313[result1313[,3]!=99 & !is.na(result1313[,3]),]

# keep only previous results and delete stat-level results
result1309 <- btw13_raw[,is.na(first.second.votes13)|first.second.votes13=="aaa"]
result1309 <- result1309[result1309[,3]!=99 & !is.na(result1309[,3]),]
result1309 <- result1309[,1:ncol(result1313)]

# make labels
main.labels13 <- main.labels13[!is.na(first.second.votes13)]
main.labels13[seq(5, length(main.labels13), by=2)] <- 
  main.labels13[seq(4, length(main.labels13)-1, by=2)] # adding the label for SMD and PR votes

main.labels13 <- paste(main.labels13, first.second.votes13[!is.na(first.second.votes13)])
main.labels13 <- gsub("aaa", "", main.labels13)

main.labels13 <- gsub("ä", "ae", main.labels13)
main.labels13 <- gsub("Ä", "AE", main.labels13)
main.labels13 <- gsub("ö", "oe", main.labels13)
main.labels13 <- gsub("Ö", "OE", main.labels13)
main.labels13 <- gsub("ü", "ue", main.labels13)
main.labels13 <- gsub("Ü", "UE", main.labels13)

colnames(result1313) <- colnames(result1309) <- main.labels13

# select columns
result1313 <- result1313[,c(1:23, 54, 55)]
result1309 <- result1309[,c(1:23, 54, 55)]

# fuse CDU and CSU into CDU/CSU
union1313 <- result1313[,grep("CDU",names(result1313))]
csu1313 <- result1313[,grep("CSU",names(result1313))]
union1313[is.na(union1313[,1]),] <- csu1313[is.na(union1313[,1]),]
result1313[,grep("CDU",names(result1313))] <- union1313

result1313 <- result1313[,grep("CSU",names(result1313),invert=T)]


union <- result1309[,grep("CDU",names(result1309))]
csu <- result1309[,grep("CSU",names(result1309))]
union[is.na(union[,1]),] <- csu[is.na(union[,1]),]
result1309[,grep("CDU",names(result1309))] <- union

result1309 <- result1309[,grep("CSU",names(result1309),invert=T)]

# create variable for year of data
result1313$Jahr <- 2013
result1309$Jahr <- 2013

# rename variables
colnames(result1313)[3] <- "Land"
colnames(result1313)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result1313)[12:13] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen")

colnames(result1309)[3] <- "Land"
colnames(result1309)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result1309)[12:13] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen")

# sort based on the WKR number
result1313 <- result1313[order(result1313$`Nr `),]
result1309 <- result1309[order(result1309$`Nr `),]

```

```{r Data09}
### BTW09 ###

# read in results
btw09_raw <- read.csv("Data/btw_kerg/btw2009_kerg.csv",
                      skip=8, sep=";", header = FALSE, encoding="latin1")

# read in first row of labels
main.labels09 <- read.csv("Data/btw_kerg/btw2009_kerg.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="latin1")

# read in second row of labels
first.second.votes09 <- read.csv("Data/btw_kerg/btw2009_kerg.csv",
                               skip=6, nrows=1, sep=";", header = FALSE)
first.second.votes09[1:3] <- "aaa"

# delete previous results and state-level results
result0909 <- btw09_raw[,!is.na(first.second.votes09)]
result0909 <- result0909[result0909[,3]!=99 & !is.na(result0909[,3]),]

# keep only previous results and delete stat-level results
result0905 <- btw09_raw[,is.na(first.second.votes09)|first.second.votes09=="aaa"]
result0905 <- result0905[result0905[,3]!=99 & !is.na(result0905[,3]),]
result0905 <- result0905[,1:ncol(result0909)]

# make labels
main.labels09 <- main.labels09[!is.na(first.second.votes09)]
main.labels09[seq(5, length(main.labels09), by=2)] <- 
  main.labels09[seq(4, length(main.labels09)-1, by=2)] # adding the label for SMD and PR votes

main.labels09 <- paste(main.labels09, first.second.votes09[!is.na(first.second.votes09)])
main.labels09 <- gsub("aaa", "", main.labels09)

main.labels09 <- gsub("ä", "ae", main.labels09)
main.labels09 <- gsub("Ä", "AE", main.labels09)
main.labels09 <- gsub("ö", "oe", main.labels09)
main.labels09 <- gsub("Ö", "OE", main.labels09)
main.labels09 <- gsub("ü", "ue", main.labels09)
main.labels09 <- gsub("Ü", "UE", main.labels09)

colnames(result0909) <- colnames(result0905) <- main.labels09

# select columns
result0909 <- result0909[,1:23]
result0905 <- result0905[,1:23]

# fuse CDU and CSU into CDU/CSU
union0909 <- result0909[,grep("CDU",names(result0909))]
csu0909 <- result0909[,grep("CSU",names(result0909))]
union0909[is.na(union0909[,1]),] <- csu0909[is.na(union0909[,1]),]
result0909[,grep("CDU",names(result0909))] <- union0909

result0909 <- result0909[,grep("CSU",names(result0909),invert=T)]


union0905 <- result0905[,grep("CDU",names(result0905))]
csu0905 <- result0905[,grep("CSU",names(result0905))]
union0905[is.na(union0905[,1]),] <- csu0905[is.na(union0905[,1]),]
result0905[,grep("CDU",names(result0905))] <- union0905

result0905 <- result0905[,grep("CSU",names(result0905),invert=T)]

# create variable for year of data
result0909$Jahr <- 2009
result0905$Jahr <- 2009

# rename variables
colnames(result0909)[3] <- "Land"
colnames(result0909)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result0909)[14:15] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen")

colnames(result0905)[3] <- "Land"
colnames(result0905)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result0905)[14:15] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen")

# sort based on the WKR number
result0909 <- result0909[order(result0909$`Nr `),]
result0905 <- result0905[order(result0905$`Nr `),]



```

```{r Data05}
### BTW05 ###

# read in results
btw05_raw <- read.csv("Data/btw_kerg/btw2005_kerg.csv",
                      skip=8, sep=";", header = FALSE, encoding="latin1")

# read in first row of labels
main.labels05 <- read.csv("Data/btw_kerg/btw2005_kerg.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="latin1")

# read in second row of labels
first.second.votes05 <- read.csv("Data/btw_kerg/btw2005_kerg.csv",
                               skip=6, nrows=1, sep=";", header = FALSE)
first.second.votes05[1:3] <- "aaa"

# delete previous results and state-level results
result0505 <- btw05_raw[,!is.na(first.second.votes05)]
result0505 <- result0505[result0505[,3]!=99 & !is.na(result0505[,3]),]

# keep only previous results and delete stat-level results
result0502 <- btw05_raw[,is.na(first.second.votes05)|first.second.votes05=="aaa"]
result0502 <- result0502[result0502[,3]!=99 & !is.na(result0502[,3]),]
result0502 <- result0502[,1:ncol(result0505)]

# make labels
main.labels05 <- main.labels05[!is.na(first.second.votes05)]
main.labels05[seq(5, length(main.labels05), by=2)] <- 
  main.labels05[seq(4, length(main.labels05)-1, by=2)] # adding the label for SMD and PR votes

main.labels05 <- paste(main.labels05, first.second.votes05[!is.na(first.second.votes05)])
main.labels05 <- gsub("aaa", "", main.labels05)

main.labels05 <- gsub("ä", "ae", main.labels05)
main.labels05 <- gsub("Ä", "AE", main.labels05)
main.labels05 <- gsub("ö", "oe", main.labels05)
main.labels05 <- gsub("Ö", "OE", main.labels05)
main.labels05 <- gsub("ü", "ue", main.labels05)
main.labels05 <- gsub("Ü", "UE", main.labels05)

colnames(result0505) <- colnames(result0502) <- main.labels05

# select columns
result0505 <- result0505[,1:23]
result0502 <- result0502[,1:23]

# fuse CDU and CSU into CDU/CSU
union0505 <- result0505[,grep("CDU",names(result0505))]
csu0505 <- result0505[,grep("CSU",names(result0505))]
union0505[is.na(union0505[,1]),] <- csu0505[is.na(union0505[,1]),]
result0505[,grep("CDU",names(result0505))] <- union0505

result0505 <- result0505[,grep("CSU",names(result0505),invert=T)]


union0502 <- result0502[,grep("CDU",names(result0502))]
csu0502 <- result0502[,grep("CSU",names(result0502))]
union0502[is.na(union0502[,1]),] <- csu0502[is.na(union0502[,1]),]
result0502[,grep("CDU",names(result0502))] <- union0502

result0502 <- result0502[,grep("CSU",names(result0502),invert=T)]

# create variable for year of data
result0505$Jahr <- 2005
result0502$Jahr <- 2005

# rename variables
colnames(result0505)[3] <- "Land"
colnames(result0505)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result0505)[14:21] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen")

colnames(result0502)[3] <- "Land"
colnames(result0502)[6:7] <- c("Waehlende Erststimmen","Waehlende Zweitstimmen")
colnames(result0502)[14:21] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen","GRUENE Erststimmen", "GRUENE Zweitstimmen","FDP Erststimmen","FDP Zweitstimmen","DIE LINKE Erststimmen","DIE LINKE Zweitstimmen")

# sort based on the WKR number
result0505 <- result0505[order(result0505$`Nr `),]
result0502 <- result0502[order(result0502$`Nr `),]


```

```{r Data02}
### BTW02 ###

# read in results
btw02_raw <- read.csv("Data/btw_kerg/btw2002_kerg.csv",
                      skip=8, sep=";", header = FALSE, encoding="latin1")

# read in first row of labels
main.labels02 <- read.csv("Data/btw_kerg/btw2002_kerg.csv",
                        skip=5, nrows=1, sep=";", header = FALSE,
                        colClasses = "character", encoding="latin1")

# read in second row of labels
first.second.votes02 <- read.csv("Data/btw_kerg/btw2002_kerg.csv",
                               skip=6, nrows=1, sep=";", header = FALSE)
first.second.votes02[1:5] <- "aaa"

# delete state-level results
result0202 <- btw02_raw[btw02_raw[,3]!=99 & !is.na(btw02_raw[,3]),]

# make labels
main.labels02 <- main.labels02[!is.na(first.second.votes02)]
main.labels02[seq(5, length(main.labels02), by=2)] <- 
  main.labels02[seq(4, length(main.labels02)-1, by=2)] # adding the label for SMD and PR votes

main.labels02 <- paste(main.labels02, first.second.votes02[!is.na(first.second.votes02)])
main.labels02 <- gsub("aaa", "", main.labels02)

main.labels02 <- gsub("ä", "ae", main.labels02)
main.labels02 <- gsub("Ä", "AE", main.labels02)
main.labels02 <- gsub("ö", "oe", main.labels02)
main.labels02 <- gsub("Ö", "OE", main.labels02)
main.labels02 <- gsub("ü", "ue", main.labels02)
main.labels02 <- gsub("Ü", "UE", main.labels02)
colnames(result0202) <- main.labels02

# select columns
result0202 <- result0202[,1:21]

# Change 0 to NA ... no candidacy
result0202[,10:21][result0202[,10:21]==0] <- NA


# fuse CDU and CSU into CDU/CSU
union0202 <- result0202[,grep("CDU",names(result0202))]
csu0202 <- result0202[,grep("CSU",names(result0202))]
union0202[is.na(union0202[,1])|union0202[,1]==0,] <-
         csu0202[is.na(union0202[,1])|union0202[,1]==0,]
result0202[,grep("CDU",names(result0202))] <- union0202

result0202 <- result0202[,grep("CSU",names(result0202),invert=T)]

# create variable for year of data
result0202$Jahr <- 2002

# rename variables
colnames(result0202)[1:2] <- c("Nr","Gebiet")
colnames(result0202)[12:13] <- c("CDU/CSU Erststimmen","CDU/CSU Zweitstimmen")

# sort based on the WKR number
result0202 <- result0202[order(result0202$Nr),]

# Recode WKR nummer
result0202$Land <- result0202$Land -900



```


## Pooling of individual datasets by election year

The current and previous election results are pooled separately.

```{r}
library(plyr)
data_now <- rbind.fill(result2525,result2121,result1717,result1313,result0909,result0505)
data_then <- rbind.fill(result2521,result1713,result1309,result0905,result0502)

## Save data for subsequent analyses
save(data_now, file = "Data/data_now.RData")
save(data_then, file = "Data/data_then.RData")

```


```{r}
load("Data/data_now.RData")
load("Data/data_then.RData")
```


# Add Time Dummy Variables

```{r}
make_year_dummies <- function(df, var = "Jahr", prefix = "d") {
  yrs <- sort(unique(df[[var]]))
  for (y in yrs) {
    df[[paste0(prefix, substr(y, 3, 4))]] <- as.integer(df[[var]] == y)  # d05, d09, d13, ...
  }
  df
}

data_now  <- make_year_dummies(data_now)
data_then <- make_year_dummies(data_then)

```

## Matrix der Treatments

```{r}
library(readxl)
treat.raw <- read_xlsx("Data/Treatment_WK_25.xlsx",sheet=2)
treat.raw <- treat.raw[!is.na(treat.raw$Jahr),]
treat.raw <- treat.raw[!is.na(treat.raw$Partei2),]

all.years <- c(2025,2021,2017,2013,2009,2005,2002)
all.parties <- c("CDU/CSU","SPD","AfD","FDP","Die Linke","Bündnis 90/Die Grünen", "BSW")
all.parties2 <- c("CDU/CSU","SPD","AfD","FDP","DIE LINKE","GRUENE", "BSW")
all.states <- c("SH","HH","NI","HB","NRW","HE","RP","BW","BY","SL","B","BB","MV","SN","SA","TH")

data_treat <- data_now # create a treatment matrix
colnames(result0202)[c(1:4,5)] <- colnames(data_now)[c(1:4,6)]
colnames(result0202)[c(18,19)] <- colnames(data_now)[c(20,21)]
result0202 <- result0202[,-21]

result0202[setdiff(names(data_now), names(result0202))] <- NA
data_treat <- rbind(data_treat,result0202) # a full dataframe with results from 2002 to 2025

data_treat[,grep("Erststimmen",names(data_treat))] <- 0 # empty cols
data_treat[,grep("Zweitstimmen",names(data_treat))] <- 0

# NEW 27.12.2025
unique(data_treat$Land)
#[1]   1  13   2   3   4  12  15  11   5  14   6  16   7   9
#[15]   8  10 901 913 902 903 904 912 915 911 905 914 906 916
#[29] 907 909 908 910
# => 2002 has different country coding!!
data_treat$Land <- ifelse(data_treat$Land > 900, 
                          data_treat$Land - 900, 
                          data_treat$Land)
print(unique(data_treat$Land))  # Should only show 1-16

for (i in 1:nrow(treat.raw)){
  
  # Nur die Zeilen für die BTW, die wir berücksichtigen wollen
  if (as.numeric(treat.raw$Jahr[i]) %in% all.years){
   
    # identify which federal election
    which.year <- as.numeric(treat.raw$Jahr[i])
    
    # identify which party's leader
    which.party <- which(all.parties==treat.raw$Partei2[i]) 
    
    # identify which treatment in which district
    which.wkr <- as.numeric(treat.raw$WK[i])    # <- original WK Nr. to match with results
    treat.wkr <- as.numeric(treat.raw$T_DM[i])
    # identify which treatment in which state
    which.land <- which(all.states == treat.raw$Land[i]) # e.g. row 2: NRW -> 5
    treat.land <- as.numeric(treat.raw$T_LP[i]) # e.g. row2: 5, KK who was there before as non KK
    ## From here replacing values in the treatment matrix
    
    # District (only in the case when the leader run for a district)
    if (!is.na(which.wkr)){
      data_treat[data_treat$Jahr==which.year & data_treat$Nr==which.wkr,
                 grep(all.parties2[which.party],names(data_treat))[1] ] <- treat.wkr
    }
    
    # Party list 
    if (length(which.land)>0){ # only in case when the state was identified
      data_treat[data_treat$Jahr==which.year & data_treat$Land==which.land,
                 grep(all.parties2[which.party],names(data_treat))[2]] <- treat.land
    }
  }
    
}

```
```{r}
# Validation: assigned treatments
treatment_assigned <- sapply(all.years, function(yr) {
  sapply(all.parties2, function(party) {
    sum(data_treat[data_treat$Jahr == yr, grep(party, names(data_treat))] > 0, na.rm = TRUE)
  })
})
print(treatment_assigned)
```

## All Objects in the Array

An array of absolute election results with five dimensions is created:

* **Constituencies** (1 … 299)
* **Parties** (1 … 6): C S A F L G
* **First or second vote** (1, 2)
* **Year** (1 … 6): 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

An array of eligible voters and valid votes with five dimensions is created:

* **Constituencies** (1 … 299)
* **Eligible voters, turnout, or valid votes** (1 … 3)
* **First or second vote** (1, 2)
* **Year** (1 … 6): 2025, 2021, 2017, 2013, 2009, 2005, 2002
* **Current or previous results** (1, 2)

An array of the treatment with four dimensions is created:

* **Constituencies** (1 … 299)
* **Parties** (1 … 6): C S A F L G
* **First or second vote** (1, 2)
* **Year** (1 … 6): 2025, 2021, 2017, 2013, 2009, 2005, 2002

A matrix of constituency names

A matrix of federal states

# FROM HERE OLD - NEEDS UPDATE

```{r}
all.abs.votes <- array(NA,dim=c(299,6,2,6,2))
all.abs.total <- array(NA,dim=c(299,3,2,6,2))
all.treat <- array(NA,dim=c(299,6,2,6))
all.wkr.name <- matrix(NA,nrow=299,ncol=6)
all.land <- matrix(NA,nrow=299,ncol=6)  

for (i.year in 1:6){
  for(i.now.then in 1:2){
    if (i.now.then==1){
      temp.dat <- data_now
      }else{
      temp.dat <- data_then
      }
    
     # BTW 2021 - 2005
     if (i.year<6){
       temp.dat <- temp.dat[(1+299*(i.year-1)):(299*i.year),]

     # Wahlergebnisse   
     # Erststimme
     all.abs.votes[,,1,i.year,i.now.then] <-
       as.matrix(temp.dat[,seq(12,22,by=2)])
     # Zweitstimme
     all.abs.votes[,,2,i.year,i.now.then] <-
       as.matrix(temp.dat[,seq(13,23,by=2)])

     # Wahlberechtigte, Wählende, Gültige
     # Erststimme
     all.abs.total[,,1,i.year,i.now.then] <- as.matrix(temp.dat[,c(4,6,10)])
     # Zweitstimme
     all.abs.total[,,2,i.year,i.now.then] <- as.matrix(temp.dat[,c(4,6,11)])
     
     }else{
       # for BTW 2002
       if (i.now.then==1){
         temp.dat <- result0202
         
     # Wahlergebnisse BTW2002  
     # Erststimme
     all.abs.votes[,c(1,2,4,5,6),1,i.year,i.now.then] <- 
                   as.matrix(temp.dat[,c(12,10,16,18,14)])
     # Zweitstimme
     all.abs.votes[,c(1,2,4,5,6),2,i.year,i.now.then] <-
                   as.matrix(temp.dat[,c(13,11,17,19,15)])
         
     # Wahlberechtigte, Wählende, Gültige
     # Erststimme
     all.abs.total[,,1,i.year,i.now.then] <- as.matrix(temp.dat[,c(4,5,8)])
     # Zweitstimme
     all.abs.total[,,2,i.year,i.now.then] <- as.matrix(temp.dat[,c(4,5,9)])

           }else{
         # Für BTW 2002 brauchen wir keine vorherige Ergebnisse
         next
       }
     }


     all.wkr.name[,i.year]  <- unlist(temp.dat[,2])
     all.land[,i.year]  <- unlist(temp.dat[,3])
  }
  
  temp.treat <- data_treat[(1+299*(i.year-1)):(299*i.year),]
  
  # Treatment
     # Erststimme
     all.treat[,,1,i.year] <- 
                   as.matrix(temp.treat[,seq(12,22,by=2)])
     # Zweitstimme
     all.treat[,,2,i.year] <-
                   as.matrix(temp.treat[,seq(13,23,by=2)])

}


```


Wahlkreiszuordnung für fixed-effect models


```{r}
wkr.assignment <- read_xlsx("Data/data_now_WK_Einteilung.xlsx",sheet=6)

wkr.assignment <- wkr.assignment[,paste0("Nr.",c("25","21","17","13","09","05","02"))]
wkr.assignment <- as.matrix(wkr.assignment)


all.wkr.index <- all.treat
for (i.party in 1:length(party.selection)){
  for (i.smdpr in 1:2){
    all.wkr.index[,i.party,i.smdpr,] <- wkr.assignment
  }
}

```



### Sozio-strukturelle Daten


```{r}

file.names <- dir("BTW_Strukturdaten")

list.struktur <- vector("list",length(file.names))

## How many lines to skip to obtain the header
to.skip.header <- c(3,3,3,4,8,8)
lines.header <- c(2,2,3,1,1,1)
to.skip.data <-c(5,5,7,6,9,9)


for (i in 1:length(file.names)){
  
  temp.header <- read.csv(paste0("BTW_Strukturdaten/",file.names[i]),
                                 skip=to.skip.header[i], 
                                 #nrows=lines.header[i],
                                 nrows=1,
                                 header=F,sep=";")
  
  
   temp.dat <- read.csv(paste0("BTW_Strukturdaten/",file.names[i]),
                                 skip=to.skip.data[i], header=F,
                                 sep=";")
  
  colnames(temp.dat) <- temp.header
  list.struktur[[i]] <- temp.dat
}


```





## Save all the objects

```{r}

save(data_now,
     data_then,
     data_treat,
     all.abs.votes,
     all.abs.total,
     all.treat,
     all.wkr.name,
     all.wkr.index,
     all.land,
  file = "Data/analysis_dat.RData")
     
```

















